Notes: New gaugings in 3d from three-form

### 2023-05-15 Draft_3dH3.tex
First draft for the full computation of the 3d potential.

\paragraph{}
The bosonic action of \EE exceptional field theory is invariant under the transformations~\eqref{eq:genlie} and has the expression
\begin{equation} \label{eq:ExFTaction}
	S_{\rm ExFT} = \int d^{3}x\,d^{248}Y\,\sqrt{\vert g\vert} \bigg(\widehat{R}+\frac{1}{240}\,D_{\mu}\M_{MN}D^{\mu}\M^{MN}+\mathscr{L}_{\rm int}+\frac{1}{\sqrt{\vert g\vert}}\,\mathscr{L}_{\rm CS}\bigg).
\end{equation}
$\widehat{R}$ is the \EE-covariantised Ricci scalar and the covariant derivative is defined as
\begin{equation}
	D_{\mu} = \partial_{\mu} - {\cal L}_{(A_{\mu}, B_{\mu})}.
\end{equation}
$\mathscr{L}_{\rm int}$ is a potential term depending only on internal derivatives:
\begin{equation}
	\begin{aligned}
		\mathscr{L}_{\rm int} &= \frac{1}{240}\,\M^{MN}\partial_{M}\M^{KL}\partial_{N}\M_{KL} - \frac{1}{2}\,\M^{MN}\partial_{M}\M^{KL}\partial_{L}\M_{NK} \\
		&\quad -\frac{1}{7200}\,f^{NQ}{}_{P}f^{MS}{}_{R}\,\M^{PK}\partial_{M}\M_{QK}\,\M^{RL}\partial_{N}\M_{SL}\\
		&\quad +\frac{1}{2}\,g^{-1}\partial_{M}g\,\partial_{N}\M^{MN} + \frac{1}{4}\,\M^{MN}\,g^{-2}\partial_{M}g\,\partial_{N}g + \frac{1}{4}\,\M^{MN}\,\partial_{M}g_{\mu\nu}\,\partial_{N}g^{\mu\nu}.
	\end{aligned}
\end{equation}
Finally, $\mathscr{L}_{\rm CS}$ is a Chern-Simons term required to impose the on-shell duality between scalar and vector fields:
\begin{equation}
	\begin{aligned}
		\mathscr{L}_{\rm CS} &= \frac{1}{2}\,\varepsilon^{\mu\nu\rho}\,\bigg(F_{\mu\nu}{}^{M}B_{\rho\,M}-f_{KL}{}^{N}\,\partial_{\mu}A_{\nu}{}^{K}\partial_{N}A_{\rho}{}^{L}-\frac{2}{3}\,f^{N}{}_{KL}\,\partial_{M}\partial_{N}A_{\mu}{}^{K}A_{\nu}{}^{M}A_{\rho}{}^{L}\\
		&\quad -\frac{1}{3}\,f_{MKL}\,f^{KP}{}_{Q}\,f^{LR}{}_{S}\,A_{\mu}{}^{M}\partial_{P}A_{\nu}{}^{Q}\partial_{R}A_{\rho}{}^{S}\bigg),
	\end{aligned}
\end{equation}
where $F_{\mu\nu}{}^{M}$ is the covariant field strength of $A_{\mu}{}^{M}$ (see eq. (2.26) of ref.~\cite{Hohm:2014fxa}).

@@

The Ansätze~\eqref{eq:GSSansatz} describe a consistent truncation if the following condition is satisfied:
\begin{equation} \label{eq:GenLeibniz}
	\gL_{\ff{U}_{\fl{M}}} {\cU}_{\fl{N}}{}^M = X_{\fl{M}\fl{N}}{}^{\fl{P}}\, \cU_{\fl{P}}{}^M,
\end{equation}
with $\cU_{\fl{M}}{}^M = \rho^{-1} U_{\fl{M}}{}^M$, $\ff{U}_{\fl{M}} = (\cU_{\fl{M}},\,\Sigma_{\fl{M}})$ and $X_{\fl{M}\fl{N}}{}^{\fl{P}}$ a constant tensor. This condition ensures that the factorised form of the Ansätze~\eqref{eq:GSSansatz} is preserved by the generalized Lie derivative, \textit{e.g.}
\begin{equation}
	\begin{aligned}
		D_{\mu}\M(x,Y) &= U^{\fl{M}}{}_{M}(Y)\,U^{\fl{N}}{}_{N}(Y)\,{\cal D}_{\mu}\M_{\fl{MN}}(x),\\
		{\cal L}_{(A_{\mu},B_{\nu})}A_{\mu}(x,Y) &= \rho(Y)^{-1}\,U_{\fl{M}}{}^{M}(Y)\,\llbracket\A_{\mu},\A_{\nu}\rrbracket^{\fl{M}}(x),
	\end{aligned}
\end{equation}
where
\begin{equation}
	{\cal D}_{\mu}\M_{\fl{MN}} = \partial_{\mu}\M_{\fl{MN}}-2\,\A_{\mu}{}^{\fl{P}}X_{\fl{P}(\fl{M})}{}^{\fl{Q}}\M_{\fl{N})\fl{Q}} \quad {\rm and} \quad \llbracket\A_{\mu},\A_{\nu}\rrbracket^{\fl{M}}=X_{\fl{PQ}}{}^{\fl{M}}\A_{\mu}{}^{\fl{P}}\A_{\nu}^{\fl{Q}}.
\end{equation}

The constant tensor $X_{\fl{M}\fl{N}}{}^{\fl{P}}$ plays the role of the embedding tensor of the three-dimensional gauged supergravity. It has the following expression:
\begin{equation}
	\begin{aligned}
		X_{\fl{MN}}{}^{\fl{P}}&=-\rho^{-1}\,\Gamma_{\fl{MN}}{}^{\fl{P}} + \rho^{-1}\,f^{\fl{P}}{}_{\fl{NQ}}f^{\fl{QK}}{}_{\fl{L}}\,\Gamma_{\fl{KM}}{}^{\fl{L}}-\frac{1}{60}\,\rho^{-1}\,f^{\fl{PK}}{}_{\fl{N}}f_{\fl{ML}}{}^{\fl{Q}}\,\Gamma_{\fl{KQ}}{}^{\fl{L}}\\
		&\quad-\frac{1}{2}\,\rho^{-1}\,f^{\fl{P}}{}_{\fl{NQ}}f^{\fl{QK}}{}_{\fl{M}}\,\Gamma_{\fl{RK}}{}^{\fl{R}}+\left(\delta_{\fl{M}}{}^{\fl{K}}\delta_{\fl{N}}{}^{\fl{P}}-\frac{1}{2}\,f_{\fl{M}}{}^{\fl{LK}}f_{\fl{NL}}{}^{\fl{P}}\right)\,\xi_{\fl{K}}.
	\end{aligned}
\end{equation}
Here, we defined the ${\rm E}_{8(8)}$ current $\Gamma_{\fl{MN}}{}^{\fl{P}}=U_{\fl{M}}{}^{K}U_{\fl{N}}{}^{L}\partial_{K}U^{\fl{P}}{}_{L}$ and the trombone gauging
\begin{equation}
	\xi_{\fl{M}} = 2\,U_{\fl{M}}{}^{N}\partial_{N}\rho^{-1}+\rho^{-1}\,\partial_{N}U_{\fl{M}}{}^{N}.
\end{equation}
In the following, we consider $\xi_{\fl{M}}=0$. The embedding tensor is most nicely expressed once projected on the adjoint representation
\begin{equation}
	X_{\fl{MN}} = -\frac{1}{60}\,X_{\fl{MP}}{}^{\fl{Q}}\,f_{\fl{NQ}}{}^{\fl{P}} = 2\,\rho^{-1}\,\Gamma_{(\fl{MN})} - \rho^{-1}\,\Gamma_{\fl{K}(\fl{M}}{}^{\fl{L}}\,f_{\fl{N})\fl{L}}{}^{\fl{K}},
\end{equation}
with the projection of the current $\Gamma_{\fl{MN}} = (1/60)\,\Gamma_{\fl{MP}}{}^{\fl{Q}}\,f_{\fl{NQ}}{}^{\fl{P}}$.

@@

With the generalized Scherk-Schwarz Ansätze~\eqref{eq:GSSansatz}, the action~\eqref{eq:ExFTaction} becomes
\begin{equation}
	\begin{aligned}
		S_{\rm ExFT} \underset{\rm gSS}{=} \int\d^{248}Y\,\rho^{-1}\int \d^{3}x & \bigg[\sqrt{\vert\mathring{g}\vert}\bigg(\mathring{R}+\frac{1}{240}\,{\cal D}_{\mu}\M_{\fl{MN}}{\cal D}^{\mu}\M^{\fl{MN}}-V\bigg) \\
		&-\varepsilon^{\mu\nu\rho}\,X_{\fl{MN}}\,\A_{\mu}{}^{\fl{M}} \left(\partial_{\nu}\A_{\rho}{}^{\fl{N}}-\frac{1}{3}\,\llbracket\A_{\nu},\A_{\rho}\rrbracket^{\fl{N}}\right)\bigg],
	\end{aligned}
\end{equation}
with the potential\footnote{To obtain this potential, one has to add the total derivative $-\partial_{M}\left(\sqrt{\vert g\vert}\,\partial_{N}\M^{MN}+(4/3)\,\M^{MN}\partial_{N}\sqrt{\vert g\vert}\right)$ to $\sqrt{\vert g\vert}\,\mathscr{L}_{\rm int}$.}
\begin{equation}
	V = X_{\fl{MN}}\,X_{\fl{PQ}}\left(\frac{1}{28}\,\M^{\fl{MP}}\M^{\fl{NQ}}+\frac{1}{2}\,\M^{\fl{MP}}\eta^{\fl{NQ}}+\frac{13}{28}\,\eta^{\fl{MP}}\eta^{\fl{NQ}}\right).
\end{equation}

>>>>

\paragraph{}
The bosonic action of \EE exceptional field theory is invariant under the transformations~\eqref{eq:genlie} and has the expression
\begin{equation} \label{eq:ExFTaction}
	S_{\rm ExFT} = \int d^{3}x\,d^{248}Y\,\sqrt{\vert g\vert} \bigg(\widehat{R}+\frac{1}{240}\,D_{\mu}\M_{MN}D^{\mu}\M^{MN}+\mathscr{L}_{\rm int}+\frac{1}{\sqrt{\vert g\vert}}\,\mathscr{L}_{\rm CS}\bigg).
\end{equation}
$\widehat{R}$ is the \EE-covariantised Ricci scalar and the covariant derivative is defined as
\begin{equation}
	D_{\mu} = \partial_{\mu} - {\cal L}_{(A_{\mu}, B_{\mu})}.
\end{equation}
$\mathscr{L}_{\rm int}$ is a potential term depending only on internal derivatives:
\begin{equation} \label{eq:ExFTLint}
	\begin{aligned}
		\mathscr{L}_{\rm int} &= \frac{1}{240}\,\M^{MN}\partial_{M}\M^{KL}\partial_{N}\M_{KL} - \frac{1}{2}\,\M^{MN}\partial_{M}\M^{KL}\partial_{L}\M_{NK} \\
		&\quad -\frac{1}{7200}\,f^{NQ}{}_{P}f^{MS}{}_{R}\,\M^{PK}\partial_{M}\M_{QK}\,\M^{RL}\partial_{N}\M_{SL}\\
		&\quad +\frac{1}{2}\,g^{-1}\partial_{M}g\,\partial_{N}\M^{MN} + \frac{1}{4}\,\M^{MN}\,g^{-2}\partial_{M}g\,\partial_{N}g + \frac{1}{4}\,\M^{MN}\,\partial_{M}g_{\mu\nu}\,\partial_{N}g^{\mu\nu}.
	\end{aligned}
\end{equation}
Finally, $\mathscr{L}_{\rm CS}$ is a Chern-Simons term required to impose the on-shell duality between scalar and vector fields:
\begin{equation}
	\begin{aligned}
		\mathscr{L}_{\rm CS} &= \frac{1}{2}\,\varepsilon^{\mu\nu\rho}\,\bigg(F_{\mu\nu}{}^{M}B_{\rho\,M}-f_{KL}{}^{N}\,\partial_{\mu}A_{\nu}{}^{K}\partial_{N}A_{\rho}{}^{L}-\frac{2}{3}\,f^{N}{}_{KL}\,\partial_{M}\partial_{N}A_{\mu}{}^{K}A_{\nu}{}^{M}A_{\rho}{}^{L}\\
		&\quad -\frac{1}{3}\,f_{MKL}\,f^{KP}{}_{Q}\,f^{LR}{}_{S}\,A_{\mu}{}^{M}\partial_{P}A_{\nu}{}^{Q}\partial_{R}A_{\rho}{}^{S}\bigg),
	\end{aligned}
\end{equation}
where $F_{\mu\nu}{}^{M}$ is the covariant field strength of $A_{\mu}{}^{M}$ (see eq. (2.26) of ref.~\cite{Hohm:2014fxa}).

@@

The Ansätze~\eqref{eq:GSSansatz} describe a consistent truncation if the following condition is satisfied:
\begin{equation} \label{eq:GenLeibniz}
	\gL_{\ff{U}_{\fl{M}}} {\cU}_{\fl{N}}{}^M = X_{\fl{M}\fl{N}}{}^{\fl{P}}\, \cU_{\fl{P}}{}^M,
\end{equation}
with $\cU_{\fl{M}}{}^M = \rho^{-1} U_{\fl{M}}{}^M$, $\ff{U}_{\fl{M}} = (\cU_{\fl{M}},\,\Sigma_{\fl{M}})$ and $X_{\fl{M}\fl{N}}{}^{\fl{P}}$ a constant tensor. This condition ensures that the factorised form of the Ansätze~\eqref{eq:GSSansatz} is preserved by the generalized Lie derivative, \textit{e.g.}
\begin{equation} \label{eq:GSSexample}
	\begin{aligned}
		D_{\mu}\M(x,Y) &= U^{\fl{M}}{}_{M}(Y)\,U^{\fl{N}}{}_{N}(Y)\,{\cal D}_{\mu}\M_{\fl{MN}}(x),\\
		{\cal L}_{(A_{\mu},B_{\nu})}A_{\mu}(x,Y) &= \rho(Y)^{-1}\,U_{\fl{M}}{}^{M}(Y)\,\llbracket\A_{\mu},\A_{\nu}\rrbracket^{\fl{M}}(x),
	\end{aligned}
\end{equation}
where
\begin{equation}
	{\cal D}_{\mu}\M_{\fl{MN}} = \partial_{\mu}\M_{\fl{MN}}-2\,\A_{\mu}{}^{\fl{P}}X_{\fl{P}(\fl{M})}{}^{\fl{Q}}\M_{\fl{N})\fl{Q}} \quad {\rm and} \quad \llbracket\A_{\mu},\A_{\nu}\rrbracket^{\fl{M}}=X_{\fl{PQ}}{}^{\fl{M}}\A_{\mu}{}^{\fl{P}}\A_{\nu}^{\fl{Q}}.
\end{equation}

For a given twist matrix $U_{\fl{M}}{}^{M}$ satisfying the consistency condition~\eqref{eq:GenLeibniz}, the action~\eqref{eq:ExFTaction} reduces to three-dimensional ${\cal N}=16$ gauged supergravity. The constant tensor $X_{\fl{M}\fl{N}}{}^{\fl{P}}$ plays the role of the embedding tensor of the three-dimensional gauged supergravity. From eq.~~\eqref{eq:GenLeibniz}, it has the following expression:
\begin{equation}
	\begin{aligned}
		X_{\fl{MN}}{}^{\fl{P}}&=-\rho^{-1}\,\Gamma_{\fl{MN}}{}^{\fl{P}} + \rho^{-1}\,f^{\fl{P}}{}_{\fl{NQ}}f^{\fl{QK}}{}_{\fl{L}}\,\Gamma_{\fl{KM}}{}^{\fl{L}}-\frac{1}{60}\,\rho^{-1}\,f^{\fl{PK}}{}_{\fl{N}}f_{\fl{ML}}{}^{\fl{Q}}\,\Gamma_{\fl{KQ}}{}^{\fl{L}}\\
		&\quad-\frac{1}{2}\,\rho^{-1}\,f^{\fl{P}}{}_{\fl{NQ}}f^{\fl{QK}}{}_{\fl{M}}\,\Gamma_{\fl{RK}}{}^{\fl{R}}+\left(\delta_{\fl{M}}{}^{\fl{K}}\delta_{\fl{N}}{}^{\fl{P}}-\frac{1}{2}\,f_{\fl{M}}{}^{\fl{LK}}f_{\fl{NL}}{}^{\fl{P}}\right)\,\xi_{\fl{K}}.
	\end{aligned}
\end{equation}
Here, we defined the ${\rm E}_{8(8)}$ current $\Gamma_{\fl{MN}}{}^{\fl{P}}=U_{\fl{M}}{}^{K}U_{\fl{N}}{}^{L}\partial_{K}U^{\fl{P}}{}_{L}$ and the trombone gauging
\begin{equation}
	\xi_{\fl{M}} = 2\,U_{\fl{M}}{}^{N}\partial_{N}\rho^{-1}+\rho^{-1}\,\partial_{N}U_{\fl{M}}{}^{N}.
\end{equation}
In the following, we consider $\xi_{\fl{M}}=0$. The embedding tensor is then most nicely expressed once projected on the adjoint representation \CE{(Give expression with projector on 3875?)}
\begin{equation} \label{eq:symembeddingtensor}
	X_{\fl{MN}} = -\frac{1}{60}\,X_{\fl{MP}}{}^{\fl{Q}}\,f_{\fl{NQ}}{}^{\fl{P}} = 2\,\rho^{-1}\,\Gamma_{(\fl{MN})} - \rho^{-1}\,\Gamma_{\fl{K}(\fl{M}}{}^{\fl{L}}\,f_{\fl{N})\fl{L}}{}^{\fl{K}},
\end{equation}
with the projection of the current $\Gamma_{\fl{MN}} = (1/60)\,\Gamma_{\fl{MP}}{}^{\fl{Q}}\,f_{\fl{NQ}}{}^{\fl{P}}$.

The three-dimensional action follows from inserting the Ansätze~\eqref{eq:GSSansatz} in the ExFT action~\eqref{eq:ExFTaction}. The expressions of the kinetic and Chern-Simons terms result from eq.~\eqref{eq:GSSexample}: \CE{(A work on the Ricci scalar?)}
\begin{equation}
	\begin{aligned}
		\mathscr{L}_{\rm kin} &= \frac{1}{240}\,D_{\mu}\M_{MN}D^{\mu}\M^{MN} \underset{\rm gSS}{=} \frac{\rho^{2}}{240}\,{\cal D}_{\mu}\M_{\fl{MN}}{\cal D}^{\mu}\M^{\fl{MN}}, \\
		\mathscr{L}_{\rm CS} &\underset{\rm gSS}{=} -\rho^{-1}\,\varepsilon^{\mu\nu\rho}\,X_{\fl{MN}}\,\A_{\mu}{}^{\fl{M}} \left(\partial_{\nu}\A_{\rho}{}^{\fl{N}}-\frac{1}{3}\,\llbracket\A_{\nu},\A_{\rho}\rrbracket^{\fl{N}}\right).
	\end{aligned}
\end{equation}

The case of the potential term $\mathscr{L}_{\rm int}$, leading to a potential $V$ for the three-dimensional scalars $\M_{\fl{MN}}$, is more subtle. We expect $V$ to be quadratic in the embedding tensor, which makes the identification of $X_{\fl{MN}}$ more complicated. We adopt the following strategy. First, we insert the generalized Sherk-Schwarz Ansätze~\eqref{eq:GSSansatz} into the expression of $\mathscr{L}_{\rm int}$~\eqref{eq:ExFTLint} with an additional total derivative: \CE{(Fill in with the eq. of 08/02, global $-1$ and  $a \rightarrow a/3$. Need to motivate the total derivative? Need to give some formula? As MMf=-fM? Ad invariance? Mention use of section constraint? Factor $\sqrt{\vert g\vert}$ to avoid carrying $\rho$ in $V$?)}
\begin{equation} \label{eq:Lintexpansion}
	\begin{aligned}
		\mathscr{L}_{\rm int} + \frac{a}{\sqrt{\vert g\vert}}\,\partial_{M}\Big(&\sqrt{\vert g\vert}\,\partial_{N}\M^{MN}+\frac{4}{3}\,\M^{MN}\partial_{N}\sqrt{\vert g\vert}\Big)\\
		&\underset{\rm gSS}{=} -\frac{1}{2}\,\M^{\fl{MN}}\,\Gamma_{\fl{MK}}\Gamma_{\fl{N}}{}^{\fl{K}} + \ldots
	\end{aligned}
\end{equation}
Then, we consider the most general exactly quadratic function in $X_{\fl{MN}}$ and develop it in \EE currents using eq.~\eqref{eq:symembeddingtensor}: \CE{(Sign convention between $V$ and ${\mathscr{L}}_{\rm int}$.)}
\begin{equation} \label{eq:Vexpansion}
	\begin{aligned}
		V &= X_{\fl{MN}}X_{\fl{PQ}}\left(\alpha\,\M^{\fl{MP}}\M^{\fl{NQ}}+\beta\,\M^{\fl{MP}}\eta^{\fl{NQ}}+\delta\,\eta^{\fl{MP}}\eta^{\fl{NQ}}+\gamma\,\eta^{\fl{MN}}\eta^{\fl{PQ}}\right) \\
		&\underset{\rm gSS}{=} \rho^{-2}\bigg(14\,\alpha\,\Gamma_{\fl{MN}}\Gamma_{\fl{PQ}}\,\M^{\fl{MP}}\M^{\fl{NQ}}+\ldots\bigg).
	\end{aligned}
\end{equation}
\CE{Finally, comparing eq.~\eqref{eq:Lintexpansion} and~\eqref{eq:Vexpansion}, the parameters are fixed to...}

@@

With the generalized Scherk-Schwarz Ansätze~\eqref{eq:GSSansatz}, the action~\eqref{eq:ExFTaction} becomes
\begin{equation}
	\begin{aligned}
		S_{\rm ExFT} \underset{\rm gSS}{=} \int\d^{248}Y\,\rho^{-1}\int \d^{3}x & \bigg[\sqrt{\vert\mathring{g}\vert}\bigg(\mathring{R}+\frac{1}{240}\,{\cal D}_{\mu}\M_{\fl{MN}}{\cal D}^{\mu}\M^{\fl{MN}}-V\bigg) \\
		&-\varepsilon^{\mu\nu\rho}\,X_{\fl{MN}}\,\A_{\mu}{}^{\fl{M}} \left(\partial_{\nu}\A_{\rho}{}^{\fl{N}}-\frac{1}{3}\,\llbracket\A_{\nu},\A_{\rho}\rrbracket^{\fl{N}}\right)\bigg],
	\end{aligned}
\end{equation}
with the potential\footnote{To obtain this potential, one has to add the total derivative $-\partial_{M}\left(\sqrt{\vert g\vert}\,\partial_{N}\M^{MN}+(4/3)\,\M^{MN}\partial_{N}\sqrt{\vert g\vert}\right)$ to $\sqrt{\vert g\vert}\,\mathscr{L}_{\rm int}$.}
\begin{equation}
	V = X_{\fl{MN}}\,X_{\fl{PQ}}\left(\frac{1}{28}\,\M^{\fl{MP}}\M^{\fl{NQ}}+\frac{1}{2}\,\M^{\fl{MP}}\eta^{\fl{NQ}}+\frac{13}{28}\,\eta^{\fl{MP}}\eta^{\fl{NQ}}\right).
\end{equation}
\CE{(A word on the 3d linear and quadratic constraints?)}
